Convective heat transport in stratified atmospheres at low and high Mach number

### 2016-10-14 tex/nusselt_paper.tex
comments out a couple of unecessary sections; revamps lots of sentences all over

\begin{abstract}
Here we study stratified convection in the context of 
plane-parallel polytropically stratified atmospheres. 
We hold the density stratification (\nrho) and Prandtl number (Pr) constant while varying the
Mach number (Ma) and the Rayleigh number (Ra) to determine the behavior of the Nusselt number (Nu), 
which quantifies the efficiency of convective heat transport.
As Ra increases and $\text{Ma} \rightarrow 1$, a scaling of Nu $\appropto$ Ra$^{0.45}$ is observed.  
As Ra increases to a regime where Ma $\geq 1$,
this scaling gives way to a weaker Nu $\appropto$ Ra$^{0.20}$.  In the regime of Ma $\ll 1$, a consistent
Nu $\appropto$ Ra$^{0.3}$ is retrieved,  reminiscent of the Nu $\propto$ Ra$^{2/7}$ seen in \RB convection.
\end{abstract}
\maketitle

@@

In this letter we study the behavior of convective heat transport, quantified by
the Nusselt number (Nu), in plane-parallel two-dimensional polytropically stratified atmospheres.  
We vary $\epsilon$ and Ra while holding \nrho and Pr
constant across all simulations.  We study domains with an aspect ratio of 4.  In section 
\ref{sec:experiment}, we describe the construction of atmospheres, our equations, and our method for
numerical time evolution.  We describe our findings in section \ref{sec:results} and discuss
implications in section \ref{sec:discussion}.

@@

Atmospheric diffusivities are specified by the Rayleigh number and the Prandtl number.  The
non-dimensional Rayleigh number is
\begin{equation}
\text{Ra} = \frac{g L_z^3 (\Delta S_0 / c_P)}{\nu\chi},
\end{equation}
where $\Delta S_0$ is the entropy difference between the top and bottom of the atmosphere, 
$c_P = R\gamma(\gamma-1)^{-1}$ is the specific heat at constant pressure,
$\nu$ is the kinematic viscosity (viscous diffusivity), and $\chi$ is the thermal diffusivity.  
The relationship between the thermal and viscous diffusivities is
set by the Prandtl number, Pr$ = \nu/\chi$.   We relate the dynamic viscosity, $\mu$, and the thermal conductivity,
$\kappa$, to their corresponding diffusivities such that 
$\nu \equiv \mu/\rho$ and $\chi \equiv \kappa/\rho$.  As a result, $\text{Ra} \propto (\nu\chi)^{-1} \propto
\rho^2$.  The atmospheres studied here with $n_{\rho} = 3$ experience an increase in the Rayleigh number 
by a factor of 400 across the domain.  This formulation leaves Pr
constant throughout the depth of the atmosphere, and in this letter we impose $\text{Pr} = 1$.
Throughout this letter, we specify Ra at the top of the atmosphere.

@@

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/ma_v_eps.png}
\caption{(a) Characteristic horizontally averaged maximum Mach numbers have been
time averaged for over $\geq 100t_{b}$ beginning roughly 50$t_{b}$ after the start of simulations. For $\epsilon \leq 0.1$,
an evident scaling of $Ma \propto \{\epsilon^{0.50}, \epsilon^{0.51}\}$ 
at Ra = $\{10^2, 10^5\}$ is retrieved.
When $\epsilon \rightarrow m_{ad}$, large deviations from this power law are seen and the systems display
supersonic dynamics.  
(b) At high epsilon, Ma scales as Ra$^{0.28}$ until it reaches the supersonic regime, at which point it turns
over and follows a power law of Ra$^{-0.10}$.  At low epsilon, consistent power laws are achieved throughout all
values of $Ra$ studied, where Ma $\propto$ \{Ra$^{0.26}$, Ra$^{0.22}$\} for $\epsilon = \{10^{-4}, 10^{-7}\}$. All
error bars are negligible.  \label{fig:ma_v_eps} }
\end{figure}

We evolve the Fully Compressible Navier-Stokes equations,
which take the form:
\begin{align}
&\begin{aligned}
&\frac{D \ln\rho}{D t} = -\Div{\bm{u}}
	\label{eqn:continuity_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{D\bm{u}}{D t}=
&-\grad T - T\grad\ln\rho + \bm{g} \\
&\,\,\,\,+\nu\Div{\lilstressT} + \lilstressT\cdot\grad\nu + \nu\lilstressT\cdot\grad\ln\rho
\label{eqn:momentum_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{D T}{D t} = -&(\gamma-1)T\Div{\bm{u}} + \frac{1}{c_V}\left( \chi\grad^2 T + \grad T \cdot \grad\chi  \right. \\ \
&+ \chi\grad T \cdot\grad\ln\rho +\nu\left[\lilstressT\cdot\nabla\right]\cdot\bm{u}\left.\right) 
	\label{eqn:energy_eqn}
\end{aligned}
\end{align}
where $D/Dt \equiv \partial_t + \bm{u}\cdot\grad$ and the viscous stress tensor is defined as
\begin{equation}
\sigma_{ij} \equiv \left(\frac{\partial u_i}{\partial x_j} + \frac{\partial u_j}{\partial x_i} - \frac{2}{3}\delta_{ij}\Div{\bm{u}}\right).
	\label{eqn:stress_tensor}
\end{equation}

In such stratified systems, the total convective flux is
\begin{equation}
\bm{F}_{\text{conv}} \equiv \bm{F}_{\text{enth}} + \bm{F}_{\text{KE}} + \bm{F}_{\text{PE}} + \bm{F}_{\text{visc}},
\end{equation}
where $\bm{F}_{\text{enth}} \equiv \rho\bm{u}(c_V T + P/\rho)$ is the enthalpy flux, $\bm{F}_{\text{KE}} \equiv 
\rho|\bm{u}|^2\bm{u}/2$ is the kinetic energy flux, $\bm{F}_{\text{PE}} \equiv \rho\bm{u}\phi$ is the potential
energy flux (with $\phi \equiv -gz$), 
and $\bm{F}_{\text{visc}} \equiv -\rho\nu\bm{u}\cdot\lilstressT$ is the viscous flux.  Taking an inner product of
Eq. \ref{eqn:momentum_eqn} with $\bm{u}$ and adding it to 
Eq. \ref{eqn:energy_eqn}, the full energy equation in conservation form is retrieved,
\begin{equation}
\frac{\partial}{\partial t}\left(\rho\left[\frac{|\bm{u}|^2}{2} + c_V T + \phi\right]\right) +
\Div{\bm{F}_{\text{conv}} + \bm{F}_{\text{cond}}} = 0
	\label{eqn:energy_eqn_full}
\end{equation}
where $\bm{F}_{\text{cond}} = -\kappa \grad T$.  An understanding of the flux terms is essential to characterizing
the convective heat transport in our systems.

The atmosphere is contained between two impenetrable, stress free, fixed temperature boundaries at
the top and bottom of the domain such that $w = \partial_z u = T_1 = 0$ at $z = \{0, L_z\}$. The domain
is periodic in the horizontal. We utilize the novel Dedalus\footnote{http://dedalus-project.org/} pseudospectral framework 
 to time-evolve Eqs. 
\ref{eqn:continuity_eqn}-\ref{eqn:energy_eqn} using an implicit-explicit third-order four-step 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}.  
Variables are time-evolved on a dealiased Chebyshev (vertical)
and Fourier (horizontal) domain in which the
physical grid dimensions are 3/2 the size of the coefficient grid.  Our physical grid sizes range from
96x384 grid points at the lowest values of Ra to 1152x4608 grid points at Ra $\geq 10^{7}$. 
By using IMEX timestepping, we implicitly step the stiff linear acoustic wave contribution and we are able to
efficiently study flows at moderate ($Ma \approx 1$) and very low ($Ma \approx 10^{-4}$)
Mach number (Fig. \ref{fig:ma_v_eps}b).  Our equations take the form
of the FC equations in \cite{lecoanet&all2014}, extended to include variable
$\nu$ and $\chi$, and we follow the approach there; this IMEX approach has been successfully 
tested against a nonlinear benchmark  of the compressible Kelvin-Helmholtz instability \cite{Lecoanet_et_al_2016_KH}.

@@

The efficiency of convection is quantified by the Nusselt number.  
While the Nusselt number is well-defined in RB convection
as the amount of total flux divided by the steady-state background conductive flux 
\cite{johnston&doering2009, otero&all2002},
a well-defined Nusselt number is more elusive in stratified convection.  A traditional definition of the Nusselt
number in stratified convection is \cite{graham1975,hurlburt&all1984}
\begin{equation}
\text{Nu} \equiv \frac{F_{\text{conv, z}} + F_{\text{cond, z}} - F_A}{F_{\text{ref}} - F_A},
\label{eqn:nusselt}
\end{equation}
where $F_{\text{conv, z}}$ and $F_{\text{cond, z}}$ are the z-components of $\bm{F}_{\text{conv}}$ and $\bm{F}_{\text{cond}}$,
respectively.  $F_A$ is the adiabatic conductive flux, defined as $F_A = -\kappa \partial_z T_{\text{ad}}$.  For an
ideal gas in hydrostatic equilibrium, $\partial_z T_{\text{ad}} \equiv - g / c_{P}$.
$F_{\text{ref}} = \Delta T / L_z$ , where $\Delta T = T(L_z) - T(0)$, is the conductive flux of a linear profile connecting the upper
and lower plates, which is constant for the choice of fixed-temperature boundaries.

We contend that this is the general form of the Nusselt number.  To illustrate this, we consider a few limiting
cases. Convection works to
suppress entropy stratification and create isentropic atmospheres.  Under the Boussinesq approximation,
density variations are ignored and entropy stratification is directly proportional to temperature stratification,
such that $\grad S \rightarrow 0$ only when $\grad T \rightarrow 0$.  Thus, for RB convection, 
$\grad T_{\text{ad}} \equiv 0$ and the familiar form of Nu is retrieved.  In the case of stratified convection,
as $\epsilon \rightarrow m_{ad} + 1$, $\grad P \propto g \rightarrow 0$ and
the resulting $\grad T_{\text{ad}} \rightarrow 0$.  In such a case, $F_A \rightarrow 0$ and the
definition of the RB Nu is appropriate to use, as convection carries all of the
flux \cite{brandenburg&all2005}. As $\epsilon \rightarrow 0$, 
$\grad T_{\text{ad}}\rightarrow \grad T_0$, the atmospheric initial temperature profile.  This causes increasingly
smaller velocity, and thermodynamic perturbations (e.g. Fig. \ref{fig:ma_v_eps}), but the removal of the 
O(1) background term in the numerator and denominator of Eq. \ref{eqn:nusselt}, makes the numerator and
denominator both O($\epsilon$).

@@

We solved initial value problems which start in hydrostatic and thermal equilibrium and experienced infinitesimal 
kicks compared to $\epsilon$ in $T_1$.  Solutions were time-evolved until a long-time average of Nu showed little
dependence on depth. By performing a linear stability analysis, we determined that the onset of convection
occurs at Ra$_c$ $= \{10.06, 10.97, 10.97\}$ for $\epsilon = \{0.5, 10^{-4}, 10^{-7}\}$, respectively.  We study Rayleigh
numbers from values at onset up to nearly $10^6$Ra$_c$ for $\epsilon = 0.5$, $10^5$Ra$_c$ for $\epsilon = 10^{-4}$,
and $10^4$Ra$_c$ for $\epsilon = 10^{-7}$.  At
high values of $\epsilon$, shock systems form in the upper atmosphere near downflow lanes 
(Fig. \ref{fig:entropy_snapshots}a) and propagate towards upflow lanes.  
These shocks heat downflowing material which reduces the efficiency of convective transport.
Such systems were reported in
both two \cite{cattaneo&all1990} and three \cite{malagoli&all1990} dimensional polytropic simulations previously.
These shocks heat material entering the downflows, reducing both the
buoyant driving and the efficiency of convective transport.

Low mach number flows, such as those in an $\epsilon = 10^{-4}$ atmosphere (Fig. \ref{fig:entropy_snapshots}b)
have similar bulk thermodynamic structure but lack the complicating dynamics of shock heating. This is interesting
in light of the current literature, as low Mach number flows are in pressure equilibrium with the background and
pressure forces cannot be the cause of narrow downflow lanes and broad upflow regions, as has been suggested
\cite{hurlburt&all1984}.  As Ra is
increased to very large values (Fig. \ref{fig:entropy_snapshots}c), thermodynamic structures 
break up into small packets which traverse through the domain many
times before diffusing rather than spanning the whole domain.  
The complicated nature of high Ra dynamics, especially in the low Ma regime where
shocks are absent, has barred us from sufficiently converging any solutions in the regime of 
$\text{Ra }> 10^5\text{Ra}_c$ at low $\epsilon$.

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/fluxes_fig.png}
\caption{Time-averaged flux profiles for high (top) and low (bottom) Mach number flows at $Ra = 10^6$.  
The dashed lines correspond to the
enthalpy flux (orange, positive) and kinetic energy flux (purple, negative).  The grey dash-dot line is the
viscous flux, the blue dash-dot-dot line is the potential energy flux, 
and the green dotted line is the radiative flux with the adiabatic radiative flux removed. All
fluxes are normalized by $F_{ref} - F_A$, as in the calculation of the Nusselt number.  The solid black line is
the properly normalized sum of all the fluxes, and under this normalization its height-averaged value is the
Nusselt number.
\label{fig:flux_profiles} }
\end{figure}

At high values of Ra, the heat transport properties of the systems become increasingly complex and time-dependent.
Large $\epsilon$ flows exhibit two local maxima in the enthalpy flux and kinetic energy flux: one in the upper 
atmosphere caused by the shocks, and one in the lower atmosphere caused by the deep mixing of convective motions
(see Fig. \ref{fig:flux_profiles}a).
At low $\epsilon$, only the deep maximum is present (Fig. \ref{fig:flux_profiles}b).  
Furthermore, the presence
of fixed-temperature boundary conditions allows the flux at the boundaries to vary.  Many runs at $Ra > 10^5$ and
$\epsilon = 10^{-4}$ exhibit time-dependent states of stratified convection (such as those shown in 
Fig. \ref{fig:entropy_snapshots}b\&c), in which the flux entering the system at the bottom of the atmosphere exceeds
that at the top.  These states are punctuated by states of vigorous shearing, similar to those previously
reported in two-dimensional RB convection \cite{goluskin&all2014}.  These shearing states flatten the bottom temperature
gradient towards adiabatic, allowing the excess energy to exit through the upper boundary.  These shearing
states will be covered in more detail in a future paper.  Regardless, a proper long-term average over shearing
and non-shearing states retrieves an invariant flux profile (and therefore Nu profile) throughout the depth
of the atmosphere.

After appropriately time-averaging the fluxes, a sensible Nusselt number is retrieved.  Nusselt numbers for
all simulations at low and high Ma are plotted in Fig. \ref{fig:nu_v_ra}.  At $\epsilon = \{10^{-4}, 10^{-7}\}$,
scaling laws of $Nu \propto Nu^{\{0.31, 0.31\}}$ are retrieved.  At $\epsilon = 0.5$, in the near-sonic
regime ($Ra \leq 10^4$), the scaling of $Nu$ with $Ra$ is inflated, with $Nu \propto Ra^{0.45}$.  As simulations
pass into the supersonic regime and shocks start to form near the downflows, propelling warm fluid deep
into the atmosphere, that scaling drops to $Nu \propto Ra^{0.19}$.  Error on all power law scalings is
negligible.

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/nu_v_ra.png}
\caption{Variation of $Nu$ as $Ra$ increases is shown for $\epsilon = \{10^{-7}, 10^{-4}, 0.5\}$. 
At high $\epsilon$, a clear transition from the subsonic to supersonic regime is evident in the scaling
of Nu with Ra.  In the low $\epsilon$ regime Nu vs. Ra scalings collapse onto a similar line which is
reminiscent of RB scalings \cite{johnston&doering2009}.  Error bars represent one standard deviation
of a long-time averaged Nu sampled throughout the depth of the domain.
\label{fig:nu_v_ra} }
\end{figure}

\section{Discussion}
\refstepcounter{section}
\label{sec:discussion}
In this letter we have studied fundamental heat transport by stratified convection in simplified 2-D polytropic
atmospheres at low and high Mach number.  Our atmospheres here are specified by two parameters, \nrho 
and $\epsilon$.  We argue that these atmospheres are the natural extension
of the RB problem to stratified systems, and should be used to understand the basic properties of stratified
convection.  The similarity between the scaling of Nu in RB convection and in these stratified polytropes suggests that
a boundary layer theory such as the Prandtl-Blausius or Grossmann-Lohse theories for incompressible flows
could be developed for fully compressible
convection in these systems \cite{ahlers&all2009}.

Furthermore, we have demonstrated that low Ma flows exhibit the same pattern of broad upflows and narrow downflows
observed in high Ma flows \cite{hurlburt&all1984}.  At low Ma, flows are often in pressure equilibrium
with their surroundings, so cooling parcels increases their density, and both of these effects reduce
entropy.  Neither of the two free thermodynamic variables can cause breaking in warm upflow regions, 
so stratification effects must cause the observed flow asymmetry.
While not studied in this work, it is possible that at low Ma, flows feel the pressure response of the
boundaries, causing deflection in a manner unseen at high Ma.

>>>>

\begin{abstract}
Here we study stratified convection in the context of 
plane-parallel, polytropically stratified atmospheres. 
We hold the density stratification (\nrho) and Prandtl number (Pr) constant while varying the
Mach number (Ma) and the Rayleigh number (Ra) to determine the behavior of the Nusselt number (Nu), 
which quantifies the efficiency of convective heat transport.
As Ra increases and $\text{Ma} \rightarrow 1$, a scaling of Nu $\propto$ Ra$^{0.45}$ is observed.  
As Ra increases to a regime where Ma $\geq 1$,
this scaling gives way to a weaker Nu $\propto$ Ra$^{0.19}$.  In the regime of Ma $\ll 1$, a consistent
Nu $\propto$ Ra$^{0.31}$ is retrieved,  reminiscent of the Nu $\propto$ Ra$^{2/7}$ seen in \RB convection.
\end{abstract}
\maketitle

@@

In this letter we study the behavior of convective heat transport, quantified by
the Nusselt number (Nu), in plane-parallel, two-dimensional, polytropically stratified atmospheres.  
$\epsilon$ and Ra are varied  while $\nrho$, Pr, and the aspect ratio are held
constant.  In section 
\ref{sec:experiment}, the construction of atmospheres, equations, and numerical methods are discussed.  
Results are described in section \ref{sec:results} and their implications are discussed
in section \ref{sec:discussion}.

@@

Atmospheric diffusivities are specified by the Rayleigh number and the Prandtl number.  The
non-dimensional Rayleigh number is
\begin{equation}
\text{Ra} = \frac{g L_z^3 (\Delta S_0 / c_P)}{\nu\chi},
\end{equation}
where $\Delta S_0$ is the entropy difference between the top and bottom of the atmosphere, 
$c_P = R\gamma(\gamma-1)^{-1}$ is the specific heat at constant pressure,
$\nu$ is the kinematic viscosity (viscous diffusivity), and $\chi$ is the thermal diffusivity.  
The relationship between the thermal and viscous diffusivities is
set by the Prandtl number, Pr$ = \nu/\chi$.   We relate the dynamic viscosity, $\mu$, and the thermal conductivity,
$\kappa$, to their corresponding diffusivities such that 
$\nu \equiv \mu/\rho$ and $\chi \equiv \kappa/\rho$.  As a result, $\text{Ra} \propto (\nu\chi)^{-1} \propto
\rho^2$.  The atmospheres studied here with $n_{\rho} = 3$ experience an increase in the Rayleigh number 
by a factor of 400 across the domain.  This formulation leaves Pr
constant throughout the depth of the atmosphere. In this letter we impose $\text{Pr} = 1$ and
specify Ra at $z = L_z$.  We study atmospheres with an aspect ratio of 4.

@@

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/ma_v_eps.png}
\caption{The maximum value of Ma which has been horizontally averaged and time averaged for $\geq 100 t_b$, 
beginning roughly $50t_b$ after
the start of simulations.  (a) For $\epsilon \leq 0.1$,
a scaling of $\text{Ma }\propto \{\epsilon^{0.50}, \epsilon^{0.51}\}$ 
at $\text{Ra }= \{10^2, 10^5\}$ exists.
When $\epsilon \rightarrow m_{ad}$, large deviations from this power law are seen.  
(b) At high $\epsilon$, Ma scales as Ra$^{0.28}$ until it reaches the supersonic regime, at which point it
follows a power law of Ra$^{-0.10}$.  At low $\epsilon$, consistent power laws are achieved throughout all
values of Ra studied, where $\text{Ma }\propto \{\text{Ra}^{0.26},\,\text{Ra}^{0.22}\}$
for $\epsilon = \{10^{-4}, 10^{-7}\}$. All
error bars are negligible.  \label{fig:ma_v_eps} }
\end{figure}

The Fully Compressible Navier-Stokes equations,
\begin{align}
&\begin{aligned}
&\frac{\partial \ln\rho}{\partial t} + \Div{\bm{u}} = -\bm{u}\cdot\grad\rho,
	\label{eqn:continuity_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial\bm{u}}{\partial t} + \grad T - &\nu\Div{\lilstressT} - \lilstressT\cdot\grad\nu = \\
&-\bm{u}\cdot\grad\bm{u} - T\grad\ln\rho + \bm{g} + \nu\lilstressT\cdot\grad\ln\rho,
\label{eqn:momentum_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial T}{\partial t} -\frac{1}{c_V}\left(\right.\chi&\left.\grad^2 T + \grad T\cdot\grad\chi\right) = \\
	&-\bm{u}\cdot\grad T - (\gamma-1)T\Div{\bm{u}} \\
	&+ \frac{1}{c_V}\left(\chi\grad T \cdot\grad\ln\rho +\nu\left[\lilstressT\cdot\nabla\right]\cdot\bm{u}\right), 
	\label{eqn:energy_eqn}
\end{aligned}
\end{align}
are evolved with a viscous stress tensor defined as
\begin{equation}
\sigma_{ij} \equiv \left(\frac{\partial u_i}{\partial x_j} + \frac{\partial u_j}{\partial x_i} - \frac{2}{3}\delta_{ij}\Div{\bm{u}}\right).
	\label{eqn:stress_tensor}
\end{equation}
These equations carry a total convective flux of
\begin{equation}
\bm{F}_{\text{conv}} \equiv \bm{F}_{\text{enth}} + \bm{F}_{\text{KE}} + \bm{F}_{\text{PE}} + \bm{F}_{\text{visc}},
\end{equation}
where $\bm{F}_{\text{enth}} \equiv \rho\bm{u}(c_V T + P/\rho)$ is the enthalpy flux, $\bm{F}_{\text{KE}} \equiv 
\rho|\bm{u}|^2\bm{u}/2$ is the kinetic energy flux, $\bm{F}_{\text{PE}} \equiv \rho\bm{u}\phi$ is the potential
energy flux (with $\phi \equiv -gz$), 
and $\bm{F}_{\text{visc}} \equiv -\rho\nu\bm{u}\cdot\lilstressT$ is the viscous flux. 
%Taking an inner product of
%Eq. \ref{eqn:momentum_eqn} with $\bm{u}$ and adding it to 
%Eq. \ref{eqn:energy_eqn}, the full energy equation in conservation form is retrieved,
%\begin{equation}
%\frac{\partial}{\partial t}\left(\rho\left[\frac{|\bm{u}|^2}{2} + c_V T + \phi\right]\right) +
%\Div{\bm{F}_{\text{conv}} + \bm{F}_{\text{cond}}} = 0,
%	\label{eqn:energy_eqn_full}
%\end{equation}
%where $\bm{F}_{\text{cond}} = -\kappa \grad T$.  
Understanding these flux terms and how they interact with the conductive flux, 
$\bm{F}_{\text{cond}} = -\kappa \grad T$, is crucial in characterizing
convective heat transport.

The atmosphere is contained between two impenetrable, stress free, fixed temperature boundaries at
the top and bottom of the domain such that $w = \partial_z u = T_1 = 0$ at $z = \{0, L_z\}$. The domain
is horizontally periodic. We utilize the novel Dedalus\footnote{http://dedalus-project.org/} pseudospectral framework 
 to time-evolve Eqs. 
\ref{eqn:continuity_eqn}-\ref{eqn:energy_eqn} using an implicit-explicit, third-order, four-step 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}.  
Variables are time-evolved on a dealiased Chebyshev (vertical)
and Fourier (horizontal) domain in which the
physical grid dimensions are 3/2 the size of the coefficient grid.  Physical grid sizes range from
96x384 grid points at the lowest values of Ra to 1152x4608 grid points at Ra $\geq 10^{7}$. 
By using IMEX timestepping, we implicitly step the stiff linear acoustic wave contribution and we are able to
efficiently study flows at moderate ($\text{Ma } \approx 1$) and very low ($\text{Ma } \approx 10^{-4}$)
Mach number (Fig. \ref{fig:ma_v_eps}b).  Our equations take the form
of the FC equations in \cite{lecoanet&all2014}, extended to include variable
$\nu$ and $\chi$, and we follow the approach there; this IMEX approach has been successfully 
tested against a nonlinear benchmark  of the compressible Kelvin-Helmholtz instability \cite{Lecoanet_et_al_2016_KH}.

@@

The efficiency of convection is quantified by the Nusselt number.  
Nu is well-defined in RB convection
as total flux normalized by the steady-state background conductive flux 
\cite{johnston&doering2009, otero&all2002}.
In stratified convection Nu is more difficulut to define, but one of the earliest definitions used
was \cite{graham1975,hurlburt&all1984}
\begin{equation}
\text{Nu} \equiv \frac{F_{\text{conv, z}} + F_{\text{cond, z}} - F_{\text{A}}}{F_{\text{ref}} - F_{\text{A}}},
\label{eqn:nusselt}
\end{equation}
where $F_{\text{conv, z}}$ and $F_{\text{cond, z}}$ are the z-components of $\bm{F}_{\text{conv}}$ and $\bm{F}_{\text{cond}}$,
respectively.  $F_{\text{A}} \equiv -\kappa \partial_z T_{\text{ad}}$ is the adiabatic conductive flux, and
$\partial_z T_{\text{ad}} \equiv - g / c_{P}$ for an ideal gas in hydrostatic equilibrium.
$F_{\text{ref}} \equiv \Delta T / L_z$, where $\Delta T \equiv T(L_z) - T(0)$ is the 
conductive flux of a linear profile connecting the upper
and lower plates, which is constant for the choice of fixed-temperature boundaries.

We contend that this is the general form of the Nusselt number and illustrate this with a few limiting
cases. Convection works to
suppress entropy stratification and create isentropic atmospheres.  Under the Boussinesq approximation,
in which density variations are ignored, entropy stratification is directly proportional to temperature stratification,
such that $\grad S \rightarrow 0$ when $\grad T \rightarrow 0$.  Thus, for RB convection, the familiar Nu is
retrieved from Eq. \ref{eqn:nusselt} because $\grad T_{\text{ad}} = 0$.  
In the case of polytropic convection,
as $\epsilon \rightarrow m_{ad} + 1$, $g \rightarrow 0$, so
$\grad T_{\text{ad}} \rightarrow 0$.  In such a case, $F_{\text{A}} \rightarrow 0$ and the
definition of the RB Nu is appropriate to use, as convection carries all of the
flux \cite{brandenburg&all2005}. As $\epsilon \rightarrow 0$, 
$\grad T_{\text{ad}}\rightarrow \grad T_0$, and increasingly
smaller velocity and thermodynamic perturbations are needed to achieve $\grad S = 0$ (Fig. \ref{fig:ma_v_eps}a).
Here, the removal of $F_{\text{A}}$ in the numerator and denominator of Eq. \ref{eqn:nusselt} makes them both
O($\epsilon$), so the overall Nu is sensible.

@@

Initial value problems were solved in which $T_1$ experienced an infinitesimal kick compared to $\epsilon$
away from hydrostatic and thermal equilibrium.
Solutions were time-evolved until a long average of Nu showed little
dependence on depth. A linear stability analysis determined that the onset of convection
occurs at $\text{Ra}_c = \{10.06, 10.97, 10.97\}$ for $\epsilon = \{0.5, 10^{-4}, 10^{-7}\}$, respectively.  Rayleigh
numbers from values at onset up to nearly $10^6$Ra$_c$ for $\epsilon = 0.5$, $10^5$Ra$_c$ for $\epsilon = 10^{-4}$,
and $10^3$Ra$_c$ for $\epsilon = 10^{-7}$ were studied.

At large $\epsilon$ (0.5), shock systems form in the upper atmosphere near downflow lanes 
(Fig. \ref{fig:entropy_snapshots}a) and propagate toward upflow lanes.  
Such systems were reported in
both two \cite{cattaneo&all1990} and three \cite{malagoli&all1990} dimensional polytropic simulations previously.
These shocks heat material entering the downflows, reducing both the
buoyant driving and the efficiency of convective transport.

Low Mach number flows, such as those in an $\epsilon = 10^{-4}$ atmosphere (Fig. \ref{fig:entropy_snapshots}b)
have similar bulk thermodynamic structure but lack the complicating dynamics of shock heating. 
Low Ma flows are in pressure equilibrium with their suroundings, so
pressure forces are unlikely to be the sole cause of symmetry breaking in up- and downflows, as has been suggested
\cite{hurlburt&all1984};  it is more likely that these flows are responding to 
the effects of density stratification.
As Ra is increased to large values (Fig. \ref{fig:entropy_snapshots}c), thermodynamic structures 
no longer span the whole domain but rather break up into small packets which traverse the domain multiple
times before diffusing.  
The complicated nature of high Ra dynamics, especially in the low Ma regime,
has prevented the convergence of solutions in the regime of 
$\text{Ra }> 10^5\text{Ra}_c$ at low $\epsilon$.

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/fluxes_fig.png}
\caption{Time-averaged flux profiles for high (top) and low (bottom) Ma flows at $\text{Ra }= 10^6$.  
The dashed lines correspond to the
enthalpy flux (orange, positive) and kinetic energy flux (purple, negative).  The grey dash-dot line is the
viscous flux, the blue dash-dot-dot line is the potential energy flux, 
and the green dotted line is the radiative flux with the adiabatic radiative flux removed. All
fluxes are normalized by $F_{\text{ref}} - F_{\text{A}}$, as in Eq. \ref{eqn:nusselt}.  The solid black line is
Nu, the properly normalized sum of all the fluxes \label{fig:flux_profiles} }
\end{figure}

At large Ra, the heat transport properties of the systems become increasingly complex and time-dependent.
Large $\epsilon$ flows exhibit two local maxima in the enthalpy flux and kinetic energy flux: one in the upper 
atmosphere caused by the shocks, and one in the lower atmosphere caused by the deep mixing of convective motions
(Fig. \ref{fig:flux_profiles}a).
At low $\epsilon$, only the deep maximum is present (Fig. \ref{fig:flux_profiles}b).  
Fixed-temperature boundary conditions allow the flux at the boundaries to vary, so many runs at $\text{Ra }> 10^5$
and $\epsilon = 10^{-4}$ exhibit states in which the flux entering the system at the bottom of the atmosphere 
exceeds that which leaves at the top.  
These states are punctuated by states of vigorous shearing, similar to those previously
reported in two-dimensional RB convection \cite{goluskin&all2014}.  These shearing states push the bottom temperature
gradient towards adiabatic, allowing the excess energy to exit through the upper boundary.  These shearing
states will be covered in more detail in a future paper.  Regardless, a proper long-term average over shearing
and non-shearing states retrieves an invariant flux (and Nu) profile throughout the depth
of the atmosphere.

After appropriately time-averaging the fluxes, a sensible Nu is retrieved.  Nusselt numbers for
all simulations at low and high Ma are plotted in Fig. \ref{fig:nu_v_ra}.  At $\epsilon = \{10^{-4}, 10^{-7}\}$,
scaling laws of $Nu \propto Nu^{\{0.31, 0.31\}}$ are retrieved.  At $\epsilon = 0.5$, in the near-sonic
regime ($Ra \leq 10^4$), the scaling of $Nu$ with $Ra$ is inflated, with $Nu \propto Ra^{0.45}$.  As simulations
pass into the supersonic regime and shocks start to form near the downflows,
that scaling drops to $Nu \propto Ra^{0.19}$.  Error on all power law scalings is negligible.

\begin{figure}[t]
\includegraphics[width=3.4375in]{./figs/nu_v_ra.png}
\caption{Variation of Nu as Ra increases at low and high Ma. 
At high $\epsilon$, a clear transition from the subsonic to supersonic regime is evident in the scaling
of Nu with Ra.  In the low $\epsilon$ regime, Nu scalings collapse onto a similar line which is
reminiscent of RB scalings \cite{johnston&doering2009}.  Error bars represent the square root of
the variance of Nu with depth.
\label{fig:nu_v_ra} }
\end{figure}

\section{Discussion}
\refstepcounter{section}
\label{sec:discussion}
In this letter we have studied fundamental heat transport by stratified convection in simplified 2-D polytropic
atmospheres which are specified by two parameters, $\nrho$ 
and $\epsilon$.  We argue that these atmospheres are the natural extension
of the RB problem to stratified systems, and should be used to understand the basic properties of stratified
convection.  The similarity between the scaling of Nu in RB convection and in low-$\epsilon$ polytropes suggests 
that a boundary layer theory such as the Grossmann-Lohse theory for incompressible flows
could be developed for fully compressible convection in these systems \cite{ahlers&all2009}.

%Furthermore, we have demonstrated that low Ma flows exhibit the same pattern of broad upflows and narrow downflows
%observed in high Ma flows \cite{hurlburt&all1984}.  At low Ma, flows are often in pressure equilibrium
%with their surroundings, so cooling parcels increases their density, and both of these effects reduce
%entropy.  Neither of the two free thermodynamic variables can cause breaking in warm upflow regions, 
%so stratification effects must cause the observed flow asymmetry.
%While not studied in this work, it is possible that at low Ma, flows feel the pressure response of the
%boundaries, causing deflection in a manner unseen at high Ma.

The dynamics of these polytropic solutions are complex and highly time-dependent, even in two dimensions.
Time-dependent oscillating shear states have developed spontaneously, as seen before in RB convection
\cite{goluskin&all2014}.  While computationally difficult, the highest values of Ra and the lowest value
of $\epsilon$ studied here are far from values found in nature.  If the scalings of Nu and Ma
presented here (Figs. \ref{fig:ma_v_eps} \& \ref{fig:nu_v_ra}) hold, then under solar conditions ($\text{Ra }\approx 10^{20}$, $\text{Ma }\approx 10^{-4}$), we expect that $\epsilon \approx 10^{-20}$ and
$\text{Nu }\approx 10^{6}$.  
Solar conditions are of course more complicated, as there $\kappa$ is
set by the radiative opacity, which depends on both $\rho$ and $T$.

Future work will aim to better understand the mechanisms of shearing states and
whether or not these states are attainable in three-dimensional, non-rotating atmospheres.  Our studies
here have set the groundwork for understanding and comparing heat transport in stratified convection
to that in RB convection \cite{johnston&doering2009}, and for future studies of transport in stratified
convection in more realistic systems, such as those bounded by stable regions \cite{hurlburt&all1986} or 
using more realistic profiles of $\kappa$.
